---
title: "Confirmatory IRT Knowledge V4 for thesis"
author: "Firdaus"
format: 
  html:
    theme: cosmo
    toc: true
    toc-location: left
    toc-depth: 6
    number-sections: true
    self-contained: true
editor: visual
---

This analysis uses thresholds: Discrimination a ≥ 0.65 (moderate) and Difficulty −3 ≤ b ≤ +3 (Baker, 2001).

# Prepare Environment

## Load Libraries

```{r warning=FALSE}
library(psych)        # descriptives
library(foreign)
library(ltm)          # 2PL via ltm()
library(irtoys)       # est(), trf()
library(mirt)         # modern IRT, M2, reliabilities
library(latticeExtra)
library(tidyverse)
library(haven)
library(writexl)
library(readxl)
library(mokken)
library(dplyr)
library(corrplot)

select <- dplyr::select
mutate <- dplyr::mutate

```

## Load Data

```{r}
data1 <- read_xlsx("CFA Knowledge V4 for thesis.xlsx")  # read data from Excel 
names(data1)                                  # List variables
dim(data1)                                    # Expect 37 items x 177 parents

```

### Recode Data

Assumes the response options are exactly: "Ya", "Tidak", "Tidak pasti" (case/spacing handled). Reverse-keyed items: K2, K3, K4, K5, K9, K11.

```{r}
# Reverse-coded items (as in your script)
reverse_items <- c("K2", "K3", "K4", "K5", "K9", "K11")

# FIX: ensure we're only recoding item columns (K1:K37) and handle case safely
item_names <- c(
  "K1","K2","K3","K4","K5","K6","K8","K9","K10","K11","K13","K14","K15",
  "K16","K17","K18","K20","K21","K22","K23","K24","K25","K26","K27","K28",
  "K29","K30","K31","K32","K33","K34","K36","K37","K38","K40","K41"
)

data2 <- data1 %>%
  mutate(across(
    .cols = setdiff(item_names, reverse_items), 
    .fns  = ~ case_when(
      tolower(as.character(.)) == "ya"           ~ 1,
      tolower(as.character(.)) == "tidak"        ~ 0,
      tolower(as.character(.)) == "tidak pasti"  ~ 2,
      TRUE ~ NA_real_
    )
  )) %>%
  mutate(across(
    .cols = reverse_items,
    .fns  = ~ case_when(
      tolower(as.character(.)) == "ya"           ~ 0,
      tolower(as.character(.)) == "tidak"        ~ 1,
      tolower(as.character(.)) == "tidak pasti"  ~ 2,
      TRUE ~ NA_real_
    )
  ))


```

### Recode to Dichotomous

```{r}
# Recode to dichotomous: 1 = correct; 0 or 2 = incorrect
data3 <- data2 %>%
  mutate(across(
    .cols = all_of(item_names),
    .fns  = ~ case_when(
      . == 1 ~ 1,
      . %in% c(0, 2) ~ 0,
      TRUE ~ NA_real_
    )
  )) %>%
   dplyr::select(all_of(item_names))   # keep item columns only

summary(data3)

```

### Create Subset Following Theoretical Domain

```{r}
domain1_items <- c("K1","K2","K3","K4","K5","K6")

domain2_items <- c("K8","K9","K10","K11","K13","K14")

domain3_items <- c("K15","K16","K17","K18","K20")

domain4a_items <- c("K21","K26","K27","K28")

domain4b_items <- c("K22","K23","K24","K25")

domain5a_items <- c("K29","K30","K31","K32","K33")

domain5b_items <- c("K34","K36","K37")

domain6_items <- c("K38","K40","K41")


```

```{r}
domain1 <- data3 %>% select(all_of(domain1_items))
domain2 <- data3 %>% select(all_of(domain2_items))
domain3 <- data3 %>% select(all_of(domain3_items))
domain4a <- data3 %>% select(all_of(domain4a_items))
domain4b <- data3 %>% select(all_of(domain4b_items))
domain5a <- data3 %>% select(all_of(domain5a_items))
domain5b <- data3 %>% select(all_of(domain5b_items))
domain6 <- data3 %>% select(all_of(domain6_items))

```

# Domain 1

## Descriptives

```{r}
descript(domain1)

```

### Checking Tetrachoric Correlation

```{r}
# Compute tetrachoric correlation matrix
tet.domain1 <- tetrachoric(domain1)
tet.domain1$rho   # <-- this is the correlation matrix

# Round and print nicely
round(tet.domain1$rho, 2)

```

```{r}
corrplot(tet.domain1$rho ,
         type  = "upper",      # show upper triangle only
         order = "hclust",     # hierarchical clustering order
         tl.col = "black",     # item labels in black
         tl.srt = 45,          # rotate labels
         addCoef.col = NULL,   # omit correlation text on the plot
         number.cex = 0.7)
```

### Checking Factors / Dimensionality

```{r}
fa.parallel(domain1, fa = "fa", cor = "tet", fm = "pa")  # principal axis FA
```

```{r}
fa.domain1 = fa(domain1, nfactors = 1, fm = "pa", rotate = "oblimin" , cor = "tet")

print(fa.domain1, digits = 3)  
```

## Performing IRT

```{r}
irt.domain1 <- ltm(domain1 ~ z1, IRT.param = TRUE)
coef(irt.domain1)
summary(irt.domain1)
```

### Remove K6

```{r}
domain1.1 <- domain1[, setdiff(names(domain1), "K6")]

```

```{r}
irt.domain1.1 <- ltm(domain1.1 ~ z1, IRT.param = TRUE)
coef(irt.domain1.1)
summary(irt.domain1.1)

```

### Remove K1

```{r}
domain1.2 <- domain1[, setdiff(names(domain1), c("K1", "K6"))]
```

```{r}
irt.domain1.2 <- ltm(domain1.2 ~ z1, IRT.param = TRUE)
coef(irt.domain1.2)
summary(irt.domain1.2)
```

### Graphical Presentation

```{r}
plot(irt.domain1.2, type = "ICC", legend = TRUE)  # Item Characteristic Curves
plot(irt.domain1.2, type = "IIC", legend = TRUE)  # Item Information Curves
plot(irt.domain1.2, items = 0, type = "IIC")      # Test Information Function
information(irt.domain1.2, c(-3,3))               # Test info in [-3,3]

```

```{r}
plot(trf(est(domain1.2, model = "2PL", engine = "ltm")))  # Test Characteristic Curve

```

```{r}
mirt.domain1.2 <- mirt(domain1.2, 1, itemtype = "2PL")
areainfo(mirt.domain1.2, c(-3,3))
```

## Unidimensionality Test

```{r}
descript(domain1.2)
```

```{r}
set.seed(2025)
# FIX: unidimTest on data (not on ltm object)
unidimTest(irt.domain1.2) 

```

```{r}
fa.domain1.2 = fa(domain1.2, nfactors = 1, fm = "pa", rotate = "oblimin" , cor = "tet")

print(fa.domain1.2, digits = 3) 
```

```{r}
mirt.domain1.2 <- mirt(domain1.2, 1, itemtype = "2PL")
coef(mirt.domain1.2, IRTpars = TRUE, simplify = TRUE)
areainfo(mirt.domain1.2, c(-3,3))   # test info (mirt)


```

## Fit Test

```{r}

margins(irt.domain1.2)          # Two-way margins fit (ltm)
person.fit(irt.domain1.2)       # Person-fit (ltm)

itemfit(mirt.domain1.2, fit_stats = "S_X2")


M2(mirt.domain1.2)                  # M2 (may be NA/err if too few df)
marginal_rxx(mirt.domain1.2)        # marginal reliability

theta_se.domain1.2 <- fscores(mirt.domain1.2, full.scores.SE = TRUE)
empirical_rxx(theta_se.domain1.2)
```

### Comparing Original Domain 1 Vs Post Removal K6 and K1

```{r}
# Extract AIC and BIC for both models
AIC_original <- AIC(irt.domain1)
BIC_original <- BIC(irt.domain1)

AIC_revised  <- AIC(irt.domain1.2)
BIC_revised  <- BIC(irt.domain1.2)

# Combine into a comparison table
comparison <- data.frame(
  Model = c("Domain1 (original)", "Domain1.2 (revised)"),
  AIC   = c(AIC_original, AIC_revised),
  BIC   = c(BIC_original, BIC_revised)
)

print(comparison)

```

# Domain 2

## Descriptives

```{r}
descript(domain2)

```

### Checking Tetrachoric Correlation

```{r}
# Compute tetrachoric correlation matrix
tet.domain2 <- tetrachoric(domain2)
tet.domain2$rho   # <-- this is the correlation matrix

# Round and print nicely
round(tet.domain2$rho, 2)

```

```{r}
corrplot(tet.domain2$rho ,
         type  = "upper",      # show upper triangle only
         order = "hclust",     # hierarchical clustering order
         tl.col = "black",     # item labels in black
         tl.srt = 45,          # rotate labels
         addCoef.col = NULL,   # omit correlation text on the plot
         number.cex = 0.7)
```

### Checking Factors / Dimensionality

```{r}
fa.parallel(domain2, fa = "fa", cor = "tet", fm = "pa")  # principal axis FA
```

```{r}
fa.domain2 = fa(domain2, nfactors = 1, fm = "pa", rotate = "oblimin" , cor = "tet")

print(fa.domain2, digits = 3) 
```

## Performing IRT

```{r}
irt.domain2 <- ltm(domain2 ~ z1, IRT.param = TRUE)
coef(irt.domain2)
summary(irt.domain2)
```

### Graphical Presentation

```{r}
plot(irt.domain2, type = "ICC", legend = TRUE)  # Item Characteristic Curves
plot(irt.domain2, type = "IIC", legend = TRUE)  # Item Information Curves
plot(irt.domain2, items = 0, type = "IIC")      # Test Information Function
information(irt.domain2, c(-3,3))               # Test info in [-3,3]

```

```{r}
# FIX: irtoys::est expects a 0/1 matrix/data.frame; use the current domain data
plot(trf(est(domain2, model = "2PL", engine = "ltm")))  # Test Characteristic Curve

```

```{r}
mirt.domain2 <- mirt(domain2, 1, itemtype = "2PL")
areainfo(mirt.domain2, c(-3,3))
```

## Unidimensionality Test

```{r}
set.seed(2025)
# FIX: unidimTest on data (not on ltm object)
unidimTest(irt.domain2) 

```

```{r}

coef(mirt.domain2, IRTpars = TRUE, simplify = TRUE)
areainfo(mirt.domain2, c(-3,3))   # test info (mirt)


```

## Fit Test

```{r}

margins(irt.domain2)          # Two-way margins fit (ltm)
person.fit(irt.domain2)       # Person-fit (ltm)

itemfit(mirt.domain2, fit_stats = "S_X2")


M2(mirt.domain2)                  # M2 (may be NA/err if too few df)
marginal_rxx(mirt.domain2)        # marginal reliability

theta_se.domain2 <- fscores(mirt.domain2, full.scores.SE = TRUE)
empirical_rxx(theta_se.domain2)
```

### Checking problem with unidimensionality

```{r}
fa.domain2 = fa(domain2, nfactors = 2, fm = "pa", rotate = "oblimin" , cor = "tet")

print(fa.domain2, digits = 3) 
```

K9 and K11 is in another factor / dimension

### Remove K11

```{r}
domain2.1 <- domain2 %>% select(-K11)
```

```{r}
descript(domain2.1)
```

```{r}
irt.domain2.1 <- ltm(domain2.1 ~ z1, IRT.param = TRUE)
coef(irt.domain2.1)
summary(irt.domain2.1)
```

```{r}
mirt.domain2.1 <- mirt(domain2.1, 1, itemtype = "2PL")
areainfo(mirt.domain2.1, c(-3,3))

coef(mirt.domain2.1, IRTpars = TRUE, simplify = TRUE)

```

#### Check unidimensionality after remove K11

```{r}
set.seed(2025)
# FIX: unidimTest on data (not on ltm object)
unidimTest(irt.domain2.1) 
```

```{r}
fa.domain2.1 = fa(domain2.1, nfactors = 1, fm = "pa", rotate = "oblimin" , cor = "tet")

print(fa.domain2.1, digits = 3) 
```

#### Graphical presentation domain 2.1

```{r}
plot(irt.domain2.1, type = "ICC", legend = TRUE)  # Item Characteristic Curves
plot(irt.domain2.1, type = "IIC", legend = TRUE)  # Item Information Curves
plot(irt.domain2.1, items = 0, type = "IIC")      # Test Information Function
information(irt.domain2.1, c(-3,3))               # Test info in [-3,3]
```

```{r}
mirt.domain2.1 <- mirt(domain2.1, 1, itemtype = "2PL")
areainfo(mirt.domain2.1, c(-3,3))
coef(mirt.domain2.1, IRTpars = TRUE, simplify = TRUE)
areainfo(mirt.domain2.1, c(-3,3))   # test info (mirt)
```

```{r}
margins(irt.domain2.1)          # Two-way margins fit (ltm)
person.fit(irt.domain2.1)       # Person-fit (ltm)

itemfit(mirt.domain2.1, fit_stats = "S_X2")


M2(mirt.domain2.1)                  # M2 (may be NA/err if too few df)
marginal_rxx(mirt.domain2.1)        # marginal reliability

theta_se.domain2.1 <- fscores(mirt.domain2.1, full.scores.SE = TRUE)
empirical_rxx(theta_se.domain2.1)
```

# Domain 3

## Descriptives

```{r}
descript(domain3)

```

### Checking Tetrachoric Correlation

```{r}
# Compute tetrachoric correlation matrix
tet.domain3 <- tetrachoric(domain3)
tet.domain3$rho   # <-- this is the correlation matrix

# Round and print nicely
round(tet.domain3$rho, 2)

```

```{r}
corrplot(tet.domain3$rho ,
         type  = "upper",      # show upper triangle only
         order = "hclust",     # hierarchical clustering order
         tl.col = "black",     # item labels in black
         tl.srt = 45,          # rotate labels
         addCoef.col = NULL,   # omit correlation text on the plot
         number.cex = 0.7)
```

### Checking Factors / Dimensionality

```{r}
fa.parallel(domain3, fa = "fa", cor = "tet", fm = "pa")  # principal axis FA
```

```{r}
fa.domain3 = fa(domain3, nfactors = 1, fm = "pa", rotate = "oblimin" , cor = "tet")

print(fa.domain3, digits = 3) 
```

## Performing IRT

```{r}
irt.domain3 <- ltm(domain3 ~ z1, IRT.param = TRUE)
coef(irt.domain3)
summary(irt.domain3)
```

### Graphical Presentation

```{r}
plot(irt.domain3, type = "ICC", legend = TRUE)  # Item Characteristic Curves
plot(irt.domain3, type = "IIC", legend = TRUE)  # Item Information Curves
plot(irt.domain3, items = 0, type = "IIC")      # Test Information Function
information(irt.domain3, c(-3,3))               # Test info in [-3,3]

```

```{r}
# FIX: irtoys::est expects a 0/1 matrix/data.frame; use the current domain data
plot(trf(est(domain3, model = "2PL", engine = "ltm")))  # Test Characteristic Curve

```

```{r}
mirt.domain3 <- mirt(domain3, 1, itemtype = "2PL")

coef(mirt.domain3, IRTpars = TRUE, simplify = TRUE)

areainfo(mirt.domain3, c(-3,3))
```

## Unidimensionality Test

```{r}
set.seed(2025)
# FIX: unidimTest on data (not on ltm object)
unidimTest(irt.domain3) 

```

```{r}

coef(mirt.domain3, IRTpars = TRUE, simplify = TRUE)
areainfo(mirt.domain3, c(-3,3))   # test info (mirt)


```

## Fit Test

```{r}

margins(irt.domain3)          # Two-way margins fit (ltm)
person.fit(irt.domain3)       # Person-fit (ltm)

itemfit(mirt.domain3, fit_stats = "S_X2")


M2(mirt.domain3)                  # M2 (may be NA/err if too few df)
marginal_rxx(mirt.domain3)        # marginal reliability

theta_se.domain3 <- fscores(mirt.domain3, full.scores.SE = TRUE)
empirical_rxx(theta_se.domain3)
```

# Domain 4a

## Descriptives

```{r}
descript(domain4a)

```

### Checking Tetrachoric Correlation

```{r}
# Compute tetrachoric correlation matrix
tet.domain4a <- tetrachoric(domain4a)
tet.domain4a$rho   # <-- this is the correlation matrix

# Round and print nicely
round(tet.domain4a$rho, 2)

```

```{r}
corrplot(tet.domain4a$rho ,
         type  = "upper",      # show upper triangle only
         order = "hclust",     # hierarchical clustering order
         tl.col = "black",     # item labels in black
         tl.srt = 45,          # rotate labels
         addCoef.col = NULL,   # omit correlation text on the plot
         number.cex = 0.7)
```

### Checking Factors / Dimensionality

```{r}
fa.parallel(domain4a, fa = "fa", cor = "tet", fm = "pa")  # principal axis FA
```

```{r}
fa.domain4a = fa(domain4a, nfactors = 1, fm = "pa", rotate = "oblimin" , cor = "tet")

print(fa.domain4a, digits = 3) 
```

## Performing IRT

```{r}
irt.domain4a <- ltm(domain4a ~ z1, IRT.param = TRUE)
coef(irt.domain4a)
summary(irt.domain4a)
```

### Graphical Presentation

```{r}
plot(irt.domain4a, type = "ICC", legend = TRUE)  # Item Characteristic Curves
plot(irt.domain4a, type = "IIC", legend = TRUE)  # Item Information Curves
plot(irt.domain4a, items = 0, type = "IIC")      # Test Information Function
information(irt.domain4a, c(-3,3))               # Test info in [-3,3]

```

```{r}
# FIX: irtoys::est expects a 0/1 matrix/data.frame; use the current domain data
#plot(trf(est(domain4a, model = "2pl", engine = "ltm")))  # Test Characteristic Curve
#Error in h(simpleError(msg, call)) :error in evaluating the argument 'x' in selecting a method for function 'plot': Lapack routine dgesv: system is exactly singular: U[1,1] = 0

```

```{r}
mirt.domain4a <- mirt(domain4a, 1, itemtype = "2PL")
areainfo(mirt.domain4a, c(-3,3))



coef(mirt.domain4a, IRTpars = TRUE, simplify = TRUE)
```

## Unidimensionality Test

```{r}
set.seed(2025)
# FIX: unidimTest on data (not on ltm object)
unidimTest(irt.domain4a) 

```

```{r}

coef(mirt.domain4a, IRTpars = TRUE, simplify = TRUE)
areainfo(mirt.domain4a, c(-3,3))   # test info (mirt)


```

## Fit Test

```{r}

margins(irt.domain4a)          # Two-way margins fit (ltm)
person.fit(irt.domain4a)       # Person-fit (ltm)

itemfit(mirt.domain4a, fit_stats = "S_X2")


M2(mirt.domain4a)                  # M2 (may be NA/err if too few df)
marginal_rxx(mirt.domain4a)        # marginal reliability

theta_se.domain4a <- fscores(mirt.domain4a, full.scores.SE = TRUE)
empirical_rxx(theta_se.domain4a)
```

# Domain 4b

## Descriptives

```{r}
descript(domain4b)

```

### Checking Tetrachoric Correlation

```{r}
# Compute tetrachoric correlation matrix
tet.domain4b <- tetrachoric(domain4b)
tet.domain4b$rho   # <-- this is the correlation matrix

# Round and print nicely
round(tet.domain4b$rho, 2)

```

```{r}
corrplot(tet.domain4b$rho ,
         type  = "upper",      # show upper triangle only
         order = "hclust",     # hierarchical clustering order
         tl.col = "black",     # item labels in black
         tl.srt = 45,          # rotate labels
         addCoef.col = NULL,   # omit correlation text on the plot
         number.cex = 0.7)
```

### Checking Factors / Dimensionality

```{r}
fa.parallel(domain4b, fa = "fa", cor = "tet", fm = "pa")  # principal axis FA
```

```{r}
fa.domain4b = fa(domain4b, nfactors = 1, fm = "pa", rotate = "oblimin" , cor = "tet")

print(fa.domain4b, digits = 3) 
```

## Performing IRT

```{r}
irt.domain4b <- ltm(domain4b ~ z1, IRT.param = TRUE)
coef(irt.domain4b)
summary(irt.domain4b)
```

### Graphical Presentation

```{r}
plot(irt.domain4b, type = "ICC", legend = TRUE)  # Item Characteristic Curves
plot(irt.domain4b, type = "IIC", legend = TRUE)  # Item Information Curves
plot(irt.domain4b, items = 0, type = "IIC")      # Test Information Function
information(irt.domain4b, c(-3,3))               # Test info in [-3,3]

```

```{r}
# FIX: irtoys::est expects a 0/1 matrix/data.frame; use the current domain data
plot(trf(est(domain4b, model = "2PL", engine = "ltm")))  # Test Characteristic Curve

```

```{r}
mirt.domain4b <- mirt(domain4b, 1, itemtype = "2PL")
areainfo(mirt.domain4b, c(-3,3))

coef(mirt.domain4b, IRTpars = TRUE, simplify = TRUE)
```

## Unidimensionality Test

```{r}
set.seed(2025)
# FIX: unidimTest on data (not on ltm object)
unidimTest(irt.domain4b) 

```

```{r}

coef(mirt.domain4b, IRTpars = TRUE, simplify = TRUE)
areainfo(mirt.domain4b, c(-3,3))   # test info (mirt)


```

## Fit Test

```{r}

margins(irt.domain4b)          # Two-way margins fit (ltm)
person.fit(irt.domain4b)       # Person-fit (ltm)

itemfit(mirt.domain4b, fit_stats = "S_X2")


M2(mirt.domain4b)                  # M2 (may be NA/err if too few df)
marginal_rxx(mirt.domain4b)        # marginal reliability

theta_se.domain4b <- fscores(mirt.domain4b, full.scores.SE = TRUE)
empirical_rxx(theta_se.domain4b)
```

# Domain 5a

## Descriptives

```{r}
descript(domain5a)

```

### Checking Tetrachoric Correlation

```{r}
# Compute tetrachoric correlation matrix
tet.domain5a <- tetrachoric(domain5a)
tet.domain5a$rho   # <-- this is the correlation matrix

# Round and print nicely
round(tet.domain5a$rho, 2)

```

```{r}
corrplot(tet.domain5a$rho ,
         type  = "upper",      # show upper triangle only
         order = "hclust",     # hierarchical clustering order
         tl.col = "black",     # item labels in black
         tl.srt = 45,          # rotate labels
         addCoef.col = NULL,   # omit correlation text on the plot
         number.cex = 0.7)
```

### Checking Factors / Dimensionality

```{r}
fa.parallel(domain5a, fa = "fa", cor = "tet", fm = "pa")  # principal axis FA
```

```{r}
fa.domain5a = fa(domain5a, nfactors = 1, fm = "pa", rotate = "oblimin" , cor = "tet")

print(fa.domain5a, digits = 3) 
```

## Performing IRT

```{r}
irt.domain5a <- ltm(domain5a ~ z1, IRT.param = TRUE)
coef(irt.domain5a)
summary(irt.domain5a)
```

### Graphical Presentation

```{r}
plot(irt.domain5a, type = "ICC", legend = TRUE)  # Item Characteristic Curves
plot(irt.domain5a, type = "IIC", legend = TRUE)  # Item Information Curves
plot(irt.domain5a, items = 0, type = "IIC")      # Test Information Function
information(irt.domain5a, c(-3,3))               # Test info in [-3,3]

```

```{r}
# FIX: irtoys::est expects a 0/1 matrix/data.frame; use the current domain data
plot(trf(est(domain5a, model = "2PL", engine = "ltm")))  # Test Characteristic Curve

```

```{r}
mirt.domain5a <- mirt(domain5a, 1, itemtype = "2PL")
areainfo(mirt.domain5a, c(-3,3))

coef(mirt.domain5a, IRTpars = TRUE, simplify = TRUE)
```

## Unidimensionality Test

```{r}
set.seed(2025)
# FIX: unidimTest on data (not on ltm object)
unidimTest(irt.domain5a) 

```

```{r}

coef(mirt.domain5a, IRTpars = TRUE, simplify = TRUE)
areainfo(mirt.domain5a, c(-3,3))   # test info (mirt)

```

## Fit Test

```{r}

margins(irt.domain5a)          # Two-way margins fit (ltm)
person.fit(irt.domain5a)       # Person-fit (ltm)

itemfit(mirt.domain5a, fit_stats = "S_X2")


M2(mirt.domain5a)                  # M2 (may be NA/err if too few df)
marginal_rxx(mirt.domain5a)        # marginal reliability

theta_se.domain5a <- fscores(mirt.domain5a, full.scores.SE = TRUE)
empirical_rxx(theta_se.domain5a)
```

# Domain 5b

## Descriptives

```{r}
descript(domain5b)

```

### Checking Tetrachoric Correlation

```{r}
# Compute tetrachoric correlation matrix
tet.domain5b <- tetrachoric(domain5b)
tet.domain5b$rho   # <-- this is the correlation matrix

# Round and print nicely
round(tet.domain5b$rho, 2)

```

```{r}
corrplot(tet.domain5b$rho ,
         type  = "upper",      # show upper triangle only
         order = "hclust",     # hierarchical clustering order
         tl.col = "black",     # item labels in black
         tl.srt = 45,          # rotate labels
         addCoef.col = NULL,   # omit correlation text on the plot
         number.cex = 0.7)
```

### Checking Factors / Dimensionality

```{r}
fa.parallel(domain5b, fa = "fa", cor = "tet", fm = "pa")  # principal axis FA
```

```{r}
fa.domain5b = fa(domain5b, nfactors = 1, fm = "pa", rotate = "oblimin" , cor = "tet")

print(fa.domain5b, digits = 3) 
```

## Performing IRT

```{r}
irt.domain5b <- ltm(domain5b ~ z1, IRT.param = TRUE)
coef(irt.domain5b)
#summary(irt.domain5b) # Error in solve.default(object$hessian) :
#Lapack routine dgesv: system is exactly singular: U[1,1] = 0
```

### Graphical Presentation

```{r}
plot(irt.domain5b, type = "ICC", legend = TRUE)  # Item Characteristic Curves
plot(irt.domain5b, type = "IIC", legend = TRUE)  # Item Information Curves
plot(irt.domain5b, items = 0, type = "IIC")      # Test Information Function
information(irt.domain5b, c(-3,3))               # Test info in [-3,3]

```

```{r}
# FIX: irtoys::est expects a 0/1 matrix/data.frame; use the current domain data
#plot(trf(est(domain5b, model = "2PL", engine = "ltm")))  # Test Characteristic Curve
#Error in h(simpleError(msg, call)) :error in evaluating the argument 'x' in selecting a method for function 'plot': Lapack routine dgesv: system is exactly singular: U[1,1] = 0

```

```{r}
mirt.domain5b <- mirt(domain5b, 1, itemtype = "2PL")
areainfo(mirt.domain5b, c(-3,3))

coef(mirt.domain5b, IRTpars = TRUE, simplify = TRUE)
```

## Unidimensionality Test

```{r}
set.seed(2025)
# FIX: unidimTest on data (not on ltm object)
unidimTest(irt.domain5b) # Error in eigen(rho., symmetric = TRUE, only.values = TRUE) :
# infinite or missing values in 'x'

```

```{r}

coef(mirt.domain5b, IRTpars = TRUE, simplify = TRUE)
areainfo(mirt.domain5b, c(-3,3))   # test info (mirt)


```

## Fit Test

```{r}

margins(irt.domain5b)          # Two-way margins fit (ltm)
person.fit(irt.domain5b)       # Person-fit (ltm)

itemfit(mirt.domain5b, fit_stats = "S_X2")


#M2(mirt.domain5b)                  # M2 (err if too few df) 
marginal_rxx(mirt.domain5b)        # marginal reliability

theta_se.domain5b <- fscores(mirt.domain5b, full.scores.SE = TRUE)
empirical_rxx(theta_se.domain5b)
```

# Domain 6

## Descriptives

```{r}
descript(domain6)

```

### Checking Tetrachoric Correlation

```{r}
# Compute tetrachoric correlation matrix
tet.domain6 <- tetrachoric(domain6)
tet.domain6$rho   # <-- this is the correlation matrix

# Round and print nicely
round(tet.domain6$rho, 2)

```

```{r}
corrplot(tet.domain6$rho ,
         type  = "upper",      # show upper triangle only
         order = "hclust",     # hierarchical clustering order
         tl.col = "black",     # item labels in black
         tl.srt = 45,          # rotate labels
         addCoef.col = NULL,   # omit correlation text on the plot
         number.cex = 0.7)
```

```{r}
irt.domain6 <- ltm(domain6 ~ z1, IRT.param = TRUE); coef(irt.domain6)


#summary(irt.domain6)

```

### Checking Factors / Dimensionality

```{r}
fa.parallel(domain6, fa = "fa", cor = "tet", fm = "pa")  # principal axis FA
```

```{r}
fa.domain6 = fa(domain6, nfactors = 1, fm = "pa", rotate = "oblimin" , cor = "tet")

print(fa.domain6, digits = 3) 
```

## Performing IRT

```{r}
irt.domain6 <- ltm(domain6 ~ z1, IRT.param = TRUE)
coef(irt.domain6)
#summary(irt.domain6)
#Error in solve.default(object$hessian) :Lapack routine dgesv: system is exactly singular: U[5,5] = 0
```

### Graphical Presentation

```{r}
plot(irt.domain6, type = "ICC", legend = TRUE)  # Item Characteristic Curves
plot(irt.domain6, type = "IIC", legend = TRUE)  # Item Information Curves
plot(irt.domain6, items = 0, type = "IIC")      # Test Information Function
information(irt.domain6, c(-3,3))               # Test info in [-3,3]

```

```{r}
# FIX: irtoys::est expects a 0/1 matrix/data.frame; use the current domain data
plot(trf(est(domain6, model = "2PL", engine = "ltm")))  # Test Characteristic Curve

```

```{r}
mirt.domain6 <- mirt(domain6, 1, itemtype = "2PL")
areainfo(mirt.domain6, c(-3,3))

coef(mirt.domain6, IRTpars = TRUE, simplify = TRUE)
```

## Unidimensionality Test

```{r}
set.seed(2025)
# FIX: unidimTest on data (not on ltm object)
unidimTest(irt.domain6) 

```

```{r}

coef(mirt.domain6, IRTpars = TRUE, simplify = TRUE)
areainfo(mirt.domain6, c(-3,3))   # test info (mirt)


```

## Fit Test

```{r}

margins(irt.domain6)          # Two-way margins fit (ltm)
person.fit(irt.domain6)       # Person-fit (ltm)

itemfit(mirt.domain6, fit_stats = "S_X2")


#M2(mirt.domain6)                  # M2 (error too few df)
marginal_rxx(mirt.domain6)        # marginal reliability

theta_se.domain6 <- fscores(mirt.domain6, full.scores.SE = TRUE)
empirical_rxx(theta_se.domain6)
```
