---
title: "final EFA Practice final 30112025 for thesis"
author: "firdaus"
format: 
  html:
    theme: cosmo
    toc: true
    toc-location: left
    toc-depth: 6
    number-sections: true
    self-contained: true
editor: visual
---

This analysis is for Practice section, with selection threshold of (FL =\>0.4), communality =\> 0.4

# Preliminaries

## Load libraries

```{r}
library(foreign)  # for importing SPSS data 
library(psych)  # for psychometrics 
library(lattice)  # easy to plot multivariate plots 
library(readxl) 
library(dplyr)
```

## Load dataset

```{r}
data1= read_xlsx("EFA_Practice_V2.xlsx")
```

```{r}
names(data1)
```

```{r}
head(data1)
```

## Coding the answers

```{r}
# Define mapping function
likert_map <- function(x) {
  recode <- c(
    "Tidak pernah" = 1,
    "Jarang"        = 2,
    "Kadang-kadang"         = 3,
    "Selalu"              = 4,
    "Sentiasa"       = 5
  )
  recode[match(tolower(x), tolower(names(recode)))]
}

# Apply transformation from data1 -> data2
data2 <- data1

for (col in names(data2)) {
  data2[[col]] <- likert_map(data2[[col]])
}

# Convert all to numeric
data2 <- as.data.frame(lapply(data2, as.numeric))

```

```{r}
head(data2)
```

```{r}
str(data2)
```

# Exploratory factor analysis

## Descriptive statistics

```{r}
describe(data2)
```

## % of response to options per item

```{r}
response.frequencies(data2)
```

## Normality of data

### Univariate normality

#### Histogram

```{r}
cat(names(data2), sep = " + ")
```

```{r}
histogram(~ P1 + P2 + P3 + P4 + P5 + P6 + P7 + P8 + P9 + P11 + P12 + P13 + P14 + P15 + P16 + P17 + P18 + P19 + P20 + P21 + P22 + P23 + P24 + P25 + P26 + P27 + P28 + P29, data = data2)
```

#### Shapiro Wilk's test

```{r}
mapply(shapiro.test, data2)
```

#### Multivariate normality

To say the data are multivariate normal:

-   *z*-kurtosis \< 5 ([Bentler, 2006](https://wnarifin.github.io/lecture/mstat/efa_medstat_practical.html#ref-bentler2006)) and the *P*-value should be ≥ 0.05

-   The plot should also form a straight line ([Arifin, 2015](https://wnarifin.github.io/lecture/mstat/efa_medstat_practical.html#ref-arifin2015))

Run Mardia’s multivariate normality test,

```{r}
mardia(data2)
```

## Step 1

### Check suitability of data for analysis

#### Kaiser-Meyer-Olkin (KMO) Measure of Sampling Adequacy (MSA)

```{r}
KMO(data2)
```

Our MSA is Marvelous (Rice,1974)

#### Bartlet's test of sphericity

```{r}
cortest.bartlett(data2)
```

### Determine the number of factors

#### Eigen value & scree plot

```{r}
scree = scree(data2)
print(scree)
```

#### Parallel analysis

```{r}
parallel = fa.parallel(data2, fm = "pa", fa = "fa")
```

#### Very simple structure (VSS) criterion & Velicer's minimum average partial (MAP) criterion)

```{r}
vss(data2)
```

## Step 2

We decided 3 factors solution after assessing 5 and 4 factors solution

### Run EFA 3 factors

```{r}
fa = fa(data2, nfactors = 3, fm = "pa", rotate = "oblimin")
print(fa, cut = .4, digits = 3)  # cut = .4 to view only FLs > 0.4
```

Factor correlation are all \<0.85

```         
      PA1   PA3   PA2
PA1 1.000 0.581 0.447
PA3 0.581 1.000 0.394
PA2 0.447 0.394 1.000
```

## Step 3

### P19 Removal

```{r}
fa1 = fa(subset(data2, select = -P19), nfactors = 3, fm = "pa", rotate = "oblimin")
print(fa1, cut = .4, digits = 3)
```

### P19, P20 Removal

```{r}
fa2 = fa(subset(data2, select = -c(P19,P20)), nfactors = 3, fm = "pa", rotate = "oblimin")
print(fa2, cut = .4, digits = 3)
```

### P19, P20, P2 Removal

```{r}
fa3 = fa(subset(data2, select = -c(P19, P20, P2)), nfactors = 3, fm = "pa", rotate = "oblimin")
print(fa3, cut = .4, digits = 3)
```

### P19, P20, P2, P3 Removal

```{r}
fa4 = fa(subset(data2, select = -c(P19, P20, P2, P3)), nfactors = 3, fm = "pa", rotate = "oblimin")
print(fa4, cut = .4, digits = 3)
```

### P19, P20, P2, P3, P4 Removal

```{r}
fa5 = fa(subset(data2, select = -c(P19, P20, P2, P3, P4)), nfactors = 3, fm = "pa", rotate = "oblimin")
print(fa5, cut = .4, digits = 3)
```

### P19, P20, P2, P3, P4, P24 Removal

```{r}
fa6 = fa(subset(data2, select = -c(P19, P20, P2, P3, P4, P24)), nfactors = 3, fm = "pa", rotate = "oblimin")
print(fa6, cut = .4, digits = 3)
```

## Summary

**PA1** : P1,P5, P6, P7, P8, P9, P11, P12, P22, P23

**PA2** : P14, P15, P16, P17, P18

**PA3** : P21, P25, P26, P27, P28, p29

# Internal Consistency Reliability

## Cronbach's Alpha

```{r}
PA1 = c("P1","P5","P6","P7","P8","P9","P11","P12","P21","P22","P23")
PA2 = c("P14","P15","P16","P17","P18")
PA3 = c("P21","P25","P26","P27","P28", "P29")
```

### PA1

```{r}
alpha.pa1 = alpha(data2[PA1]) #error unknown colour name
print(alpha.pa1, digits = 3)

#alpha.pa1 <- psych::alpha(data2[PA1], check.keys = TRUE)
#print(alpha.pa1, digits = 3)
```

### PA2

```{r}
alpha.pa2 = alpha(data2[PA2])
print(alpha.pa2, digits = 3)

#alpha.pa2 <- psych::alpha(data2[PA2], check.keys = TRUE)
#print(alpha.pa2, digits = 3)
```

### PA3

```{r}
alpha.pa3 = alpha(data2[PA3])
print(alpha.pa3, digits = 3)

#alpha.pa3 <- psych::alpha(data2[PA3], check.keys = TRUE)
#print(alpha.pa3, digits = 3)
```
