---
title: "Final EFA Attitude 30112025 for thesis"
author: "firdaus"
format: 
  html:
    theme: cosmo
    toc: true
    toc-location: left
    toc-depth: 6
    number-sections: true
    self-contained: true
editor: visual
---

This analysis is for Attitude section, with selection threshold higher (FL =\>0.4), communality =\> 0.4

# Preliminaries

## Load libraries

```{r}
library(foreign)  # for importing SPSS data
library(psych)  # for psychometrics
library(lattice)  # easy to plot multivariate plots
library(readxl)
library(dplyr)
```

## Load dataset

```{r}
data1= read_xlsx("EFA_Attitude_V2.xlsx")
```

```{r}
names(data1)
```

```{r}
head(data1)
```

## Coding the answers

```{r}
# Define mapping functions
likert_map <- function(x) {
  recode <- c(
    "Sangat tidak setuju" = 1,
    "Tidak setuju"        = 2,
    "Tidak pasti"         = 3,
    "Setuju"              = 4,
    "Sangat setuju"       = 5
  )
  recode[match(tolower(x), tolower(names(recode)))]
}

likert_reverse <- function(x) {
  recode <- c(
    "Sangat tidak setuju" = 5,
    "Tidak setuju"        = 4,
    "Tidak pasti"         = 3,
    "Setuju"              = 2,
    "Sangat setuju"       = 1
  )
  recode[match(tolower(x), tolower(names(recode)))]
}

# List of reverse-coded items
reverse_items <- c("A7","A8","A10","A11","A12","A13","A14",
                   "A24","A25","A27","A30","A32","A34")

# Apply transformation from data1 -> data2
data2 <- data1

for (col in names(data2)) {
  if (col %in% reverse_items) {
    data2[[col]] <- likert_reverse(data2[[col]])
  } else {
    data2[[col]] <- likert_map(data2[[col]])
  }
}

# Convert all to numeric
data2 <- as.data.frame(lapply(data2, as.numeric))

```

```{r}
head(data2)
```

```{r}
str(data2)
```

# Exploratory factor analysis

## Descriptive statistics

```{r}
describe(data2)
```

## % of response to options per item

```{r}
response.frequencies(data2)
```

## Normality of data

### Univariate normality

#### Histogram

```{r}
cat(names(data2), sep = " + ")
```

```{r}
histogram(~ A1 + A2 + A3 + A4 + A5 + A6 + A7 + A8 + A10 + A11 + A12 + A13 + A14 + A16 + A19 + A20 + A21 + A22 + A23 + A24 + A25 + A26 + A27 + A28 + A30 + A31 + A32 + A33 + A34 + A35,
          data = data2)
```

By eye-balling method, most of item answers were not normality distributed

#### Shapiro Wilk's test

```{r}
mapply(shapiro.test, data2)
```

All item were not normally distributed

#### Multivariate normality

To say the data are multivariate normal:

-   *z*-kurtosis \< 5 ([Bentler, 2006](https://wnarifin.github.io/lecture/mstat/efa_medstat_practical.html#ref-bentler2006)) and the *P*-value should be ≥ 0.05

-   The plot should also form a straight line ([Arifin, 2015](https://wnarifin.github.io/lecture/mstat/efa_medstat_practical.html#ref-arifin2015))

Run Mardia’s multivariate normality test,

```{r}
mardia(data2)
```

Kurtosis \< 0.001, multivariate not normally distributed

## Step 1

### Check suitability of data for analysis

#### Kaiser-Meyer-Olkin (KMO) Measure of Sampling Adequacy (MSA)

```{r}
KMO(data2)
```

Our MSA is meritorious

#### Bartlet's test of sphericity

```{r}
cortest.bartlett(data2)
```

*P*-value \< 0.05, significant. Items are correlated and worthwhile for factor analysis

### Determine the number of factors

#### Eigen value & scree plot

```{r}
scree = scree(data2)
print(scree)
```

Eigen values of factors \> 1 = 4

5 dots are above the Eigen values = 1

Thus, based on our judgement on the scree plot and eigenvalues (of factor analysis), the suitable number of factors = 4 or 5

#### Parallel analysis

```{r}
parallel = fa.parallel(data2, fm = "pa", fa = "fa")
```

There are five dots above the dashed lines. This is suggestive of 5 factors

#### Very simple structure (VSS) criterion & Velicer's minimum average partial (MAP) criterion)

```{r}
vss(data2)
```

VSS complexity of 1 indicates 3/5 factors (`vss1` largest at 1 and 2 factors) while MAP suggest 6 factors.

The suggestive factors are from 1 - 6 factors

## Step 2

We decided for 4 factors as the best after checking 5 and 3 factors

```{r}
fa = fa(data2, nfactors = 4, fm = "pa", rotate = "oblimin")
print(fa, cut = .5, digits = 3)  # cut = .3 to view only FLs > 0.3
```

### Run EFA 3 factors

Cut off factor loading based on Hair (2019) ( =\> 0.4 practically significant structure)

```{r}
fa = fa(data2, nfactors = 4, fm = "pa", rotate = "oblimin")
print(fa, cut = .4, digits = 3)  # cut = .4 to view only FLs > 0.4
```

**Poor items identified**

A27 (poor FL and communality)

A24 (poor communality)

A25 (poor communality)

A28 (poor communality)

A30 (poor communality)

A31 (poor communality)

A34 (poor communality)

No factor with cross-loading

Factor correlation is all \<0.85

```         
      PA1   PA3   PA4   PA2
PA1 1.000 0.589 0.308 0.144
PA3 0.589 1.000 0.220 0.080
PA4 0.308 0.220 1.000 0.406
PA2 0.144 0.080 0.406 1.000
```

## Step 3

### Item A27 Removal

```{r}
fa1 = fa(subset(data2, select = -A27), nfactors = 4, fm = "pa", rotate = "oblimin")
print(fa1, cut = .4, digits = 3)
```

#### Remove A27 & A24

```{r}
fa2 = fa(subset(data2, select = -c(A27, A24)), nfactors = 4, fm = "pa", rotate = "oblimin")
print(fa2, cut = .4, digits = 3)
```

#### Remove A27, A24 & A25

```{r}
fa3 = fa(subset(data2, select = -c(A27, A24, A25)), nfactors = 4, fm = "pa", rotate = "oblimin")
print(fa3, cut = .4, digits = 3)

```

#### Remove A27, A24, A25, A28

```{r}
fa4 = fa(subset(data2, select = -c(A27, A24, A25, A28)), nfactors = 4, fm = "pa", rotate = "oblimin")
print(fa4, cut = .5, digits = 3)
```

#### Remove A27, A24, A25, A28, A30

```{r}
fa5 = fa(subset(data2, select = -c(A27, A24, A25, A28, A30)), nfactors = 4, fm = "pa", rotate = "oblimin")
print(fa5, cut = .4, digits = 3)
```

#### Remove A27, A24, A25, A28, A30, A34

```{r}
fa6 = fa(subset(data2, select = -c(A27, A24, A25, A28, A30, A34)), nfactors = 4, fm = "pa", rotate = "oblimin")
print(fa6, cut = .4, digits = 3)
```

#### Remove A27, A24, A25, A28, A30, A34, A32

```{r}
fa7 = fa(subset(data2, select = -c(A27, A24, A25, A28, A30, A34, A32)), nfactors = 4, fm = "pa", rotate = "oblimin")
print(fa7, cut = .4, digits = 3)
```

#### Remove A27, A24, A25, A28, A30, A34, A32, A31

```{r}
fa8 = fa(subset(data2, select = -c(A27, A24, A25, A28, A30, A34, A32, A31)), nfactors = 4, fm = "pa", rotate = "oblimin")
print(fa8, cut = .4, digits = 3)
```

## Summary

**PA1** : A1, A2, A3, A4, A5, A6

**PA2** : A7, A8, A10

**PA3** : A11,A12,A13,A14

**PA4** : A16,A19,A20,A21,A22,A23,A26,A33,A35

# Internal Consistency Reliability

## Cronbach's Alpha

```{r}
PA1 = c("A1","A2","A3","A4","A5","A6")
PA2 = c("A7","A8","A10")
PA3 = c("A11","A12","A13","A14")
PA4 = c("A16","A19","A20","A21","A22","A23","A26","A33","A35")
```

### PA1

```{r}
alpha.pa1 = psych::alpha(data2[PA1])
print(alpha.pa1, digits = 3)
```

### PA2

```{r}
alpha.pa2 = psych::alpha(data2[PA2])
print(alpha.pa2, digits = 3)
```

### PA3

```{r}
alpha.pa3 = psych::alpha(data2[PA3])
print(alpha.pa3, digits = 3)
```

### PA4

```{r}
alpha.pa4 = psych::alpha(data2[PA4])
print(alpha.pa4, digits = 3)
```
